Applied Measure Theory for Probabilistic Modeling

### 2022-05-11 paper/paper.tex
edits

Throughout this discussion, it's important to keep in mind that for us the measure-theoretic abstractions are only means to an end.% 
\footnote{The stackoverflow discussion \url{https://mathoverflow.net/q/11591} discusses books on measure theory; M.S.~likes \cite{Shiryaev1996} as starting point and uses \cite{Elstrodt2011} and \cite{Bauer_Heinz1982-01-01} as reference. As we anticipate interest in connecting with the actual theory, we give some pointers to more technical details in the footnotes.}

Our primary interest, and the goal of \verb|MeasureTheory|, is to offer support for \emph{applied probabilistic modeling}.

The package name is owed to the fact, that the word measure alone is too vague.

Given a space $X$ of possible outcomes and a set of subsets of $X$ called \href{https://en.wikipedia.org/wiki/Event_(probability_theory)}{\emph{events}}, a probability distribution assigns each event a non-negative quantity, called the \emph{probability} or \emph{probability mass}.

Likewise, a \emph{measure} assigns each \emph{measurable set} a non-negative quantity, also called a ``mass''.

@@

One further such space that deserves mention is traditionally denoted $\Omega$. This is an abstract space connected to the real-world notion of \emph{probability}, which we denote as $\operatorname{Prob}$.

Computationally, $\Omega$ can be considered to be the set of possible states of a random number generator \verb|rng::AbstractRNG|, which is the source of computational randomness.

@@

In fact, it is possible for small enough spaces to assign meaningful probabilities to all subsets of the space, but impossible for any space containing a continuum such as the real line. Just representing a single arbitrary subset of the \verb|Float64|-range, the computational abstraction of the real line, would require staggering 2\,306 Petabytes of memory. This is not practical.

>>>>

Throughout this discussion, it's important to keep in mind that for us the measure-theoretic abstractions are only means to an end.% 
\footnote{The stackoverflow discussion \url{https://mathoverflow.net/q/11591} discusses books on measure theory; M.S.~likes \cite{Shiryaev1996} as starting point and uses \cite{Elstrodt2011} and \cite{Bauer_Heinz1982-01-01} as reference. As we anticipate interest in connecting with the actual theory, we give some pointers to more technical details in the footnotes.}
Our primary interest, and the goal of \verb|MeasureTheory|, is to offer support for \emph{applied probabilistic modeling}. 
The package name is owed to the fact, that the word measure alone is too vague.

Given a space $X$ of possible outcomes and a set of subsets of $X$ called \href{https://en.wikipedia.org/wiki/Event_(probability_theory)}{\emph{events}}, a probability distribution assigns each event a non-negative quantity, called the \emph{probability} or \emph{probability mass}.
Likewise, a \emph{measure} assigns each \emph{measurable set} a non-negative quantity, also called a ``mass''.

@@

One further such space that deserves mention is traditionally denoted $\Omega$. This is an abstract space connected to the real-world notion of \emph{probability}, which we denote as $\operatorname{Prob}$.
Computationally, $\Omega$ can be considered to be the set of possible states of a random number generator \verb|rng::AbstractRNG|, which is the source of computational randomness.

@@

For spaces that are ``small enough'', it is in fact possible to assign meaningful probabilities to all subsets of the space. Unfortunately, this is impossible for any space containing a continuum such as the real line. Just representing a single arbitrary subset of the \verb|Float64|-range, the computational abstraction of the real line, would require staggering 2\,306 Petabytes of memory. This is not practical.
